Projection variability of clausal complements across different operators

### 2024-02-05 writing/SuB-paper/projection-operators-SuB.tex
sub paper edits before overleaf

{\large \textbf{Projection variability of clausal complements across different operators}}\footnote{We would like to thank\ldots Taylor Mahler; Reviewers at SALT, SuB, XPrag; Audience at Syntax \& Semantics colloquium at University of Stuttgart}\\
Lisa HOFMANN --- \textit{University of Stuttgart}\\
Marie-Catherine DE MARNEFFE --- \textit{UC Louvain}\\
Judith TONHAUSER --- \textit{University of Stuttgart}\\

\textbf{Abstract.}
	We present experimental evidence that the projection of clausal complements \textbf{(i)} varies between entailment-canceling operators, \textbf{(ii)} that the effect of operator varies between clause-embedding predicates, and \textbf{(iii)} we extend the result from \citet{degen_are_2022}, that projection ratings in polar questions are gradient in ways that cannot be predicted by categorical lexical classes, to contexts with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
	%
	The observed variability is not captured by existing theoretical accounts of projection
	(e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}).
	%
	Our results suggest that an analysis must consider interactions between predicates and operators and raise important questions for future research on projection.

@@

\section{Introduction}
	Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC, here: \emph{Julian dances salsa,})
	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditionals \ref{ex:cond}, in which case we say that it \textit{projects} (\citealt{frege_uber_1892,strawson_referring_1950,kiparsky_fact_1970,karttunen_conventional_1979}, and many more).

@@

	We investigate the question whether the projection of embedded propositional content across entailment-cancelling operators differs by operator, and whether by-operator differences vary between embedding predicates.
	%
	Textbook diagnostics for projection often assume that, if a content projects across one of the operators in \Last, it uniformly projects across the others (\citealt{chierchia_meaning_1990}). Accordingly, current research examines projection variability between these operators only rarely, and with conflicting findings.

@@

	\citet{smith_relationship_2014} investigated by-operator variation, comparing negation and conditionals, for various types of projective content. They found that the projective content of epithets (e.g. \textit{idiot}) and the CC of \textit{know} was more projective under negation than conditionals, whereas that of appositive relative clauses and the preparatory content of \textit{win} showed the opposite pattern, and that of clefts showed no difference. Their task involved asking participants how surprised they would be to learn the content under investigation after observing the utterance. Since other factors besides speaker commitment modulate discourse (un)expectedness \citep[see e.g.][]{zimmermann_grammatical_2011,tonnis_german_2021}, it is unclear that they (only) measured projection.

	\posscite{sieker_projective_2022} study on German clause-embedding predicates compared the projection of purported factives (\textit{wissen} `know', \textit{bereuen} `regret', \textit{enthüllen} `reveal') and semi-factives (\textit{bemerken} `notice', \textit{entdecken} `discover', \textit{herausfinden} `find out') across the four contexts in \Last. They replicated Smith and Hall's result that the CC of \textit{know} projects more from under negation than conditionals for German \textit{wissen} (`know'). However, this finding was part of an overall pattern of higher projection ratings with negation than other operators. They found no interaction with predicate type, and therefore no evidence for the distinction of factives vs. semi-factives. This is based on an analysis which aggregated the ratings for the four purported factive and semi-factive predicates, under each of the entailment-cancelling operators.

	Studies in \citet{djarv_cognitive_2018}, \citet{tonhauser_how_2018}, and \citet{degen_are_2022} observed by-predicate variation in polar questions, where Karttunen's classification predicts more projection for factives than semi-factives.
	%
	\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, did find higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive). But here, it is also not obvious how exactly this task relates to projection.
	%
	\citet{tonhauser_how_2018} and \citet{degen_are_2022} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC. The observed differences between individual clause-embedding predicates do not align with theoretically assumed lexical classes. While Degen \& Tonhauser demonstrate this for the distinction between purported factive and non-factive predicates, the results also did not match the expectations from Karttunen's factive vs. semi-factive classification.
	%
	Using speaker-certainty ratings, \citet{tonhauser_how_2018} found, e.g., that the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}. \citet{degen_are_2022} observed that the CC of factive \textit{reveal} was less projective than that of semi-factive \textit{discover}, which, in turn, was less projective than that of factive \textit{know}. These findings suggest that individually examining projection ratings for various predicates, rather than pre-grouping them based on theoretical assumptions, can provide more nuanced insights into by-predicate variation in operator effects on projection.

	While \citet{smith_relationship_2014} found by-predicate variation in the effect of the operator on projection, \citet{sieker_projective_2022} did not.
	%
	These divergent results raise the question if they are due to cross-linguistic variation, task differences, or artifacts of the analysis. To address this, we conducted a series of experiments designed to assess projection across the four entailment-canceling operators in \ref{ex:family}. We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive (\emph{be annoyed, know, reveal}) and semi-factive predicates (\emph{discover, see}), and 15 non-factive predicates, given recent findings that their complements may also project, albeit to varying degrees (\citealt{degen_are_2022}).

@@

	To assess the effect of entailment-cancelling operators and clause-embedding predicates on the projection of the embedded content, we collected projectivity judgments for contents embedded under 20 clause-embedding predicates in four sets of experiments. The predicates were embedded under negation in Exps.~1, under polar questions (Exps.~2), under the epistemic possibility modal {\em perhaps} (Exps.~3), and in conditional antecedents (Exps.~4). (Each set of experiments contained three experiments differing in an at-issueness measure used in a separate block. We focus on the projection ratings here.)

		In each experiment, participants read utterances like those in \ref{ex:family} and judged whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?).

		\begin{itemize}
			\item say more about task
		\end{itemize}

		Based on Karttunen's generalization, we would expect the CC of semi-factive predicates to receive high projection ratings under negation and lower ratings under the other operators, while the CC of factive predicates would consistently receive relatively high projection ratings throughout.

@@

		\subsubsection{Participants}
			We analyze the data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform. (roughly 750 participants in each experiment)

		\subsubsection{Materials}
			Each participant saw all 20 attitude predicates (each paired with a unique content from a set of 20 contents) under one operator.

			\begin{itemize}
				\item 20 clause-embedding predicates that have shown projection variability in question contexts (Degen \& Tonhauser, 2022)
				\item Crossed w/ 20 CCs: 20 x 20 = 400 combinations
				\item One experiment per operator
				\item six main-clause controls for data exclusion
			\end{itemize}

		\subsubsection{Procedure}

			On each trial, participants read an utterance, which was presented as an utterance of a speaker with a randomly chosen name, as well as the corresponding projection question, and then gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other. This is illustrated in Fig.~\ref{fig:trial}.

			Following \citet{tonhauser_how_2018}, we took a `yes' response to indicate that participants interpret the utterance in a way that the presented speaker (e.g., Christopher in Fig.~\ref{fig:trial}) is committed to the embedded content, which reflects projection. A `no' response was taken to indicate that the content does not project.

@@

	\subsection{Results and discussion}
		Our analysis reveals two key results: \textbf{(i)} There is projection variability by operator and \textbf{(ii)} there is by-predicate variation in the effect of operator on projection.

		\subsubsection{By-operator variation}

			We find an effect of operator when aggregating across predicates. Mean projection ratings were higher under question-embeddings than under negation and modals, but lower than in conditional antecedents. This is shown in Fig.~\ref{fig:op-ratings}, and the generalizations are supported by Model \#1 reported in Table~\ref{t:op-model}.

			\begin{figure}[ht]
				\centering
				\includegraphics[width = .35\linewidth]{projective-op}
				\caption{Mean certainty ratings by operator with $95\%$ bootstrapped confidence intervals.}
				\label{fig:op-ratings}
			\end{figure}

@@

					\caption{Excerpt from the model output of a linear mixed effects model with fixed effects of operator; random effects: participant and item intercepts, fit with \texttt{lme4, lmertest} in \texttt{R}.\label{t:op-model}}
				\end{table}

			This result differs from that of \cite{sieker_projective_2022}, who observed that the CC of German attitude predicates projects more from under negation than the other operators.

			\begin{itemize}
				\item but like them, we observe small differences 
				\item overall, they have much higher projection ratings
				\item but they chose a different set of predicates, we can understand these different results when taking into account by-predicate variation in the effect of operator
			\end{itemize}

@@

			We observe differences across the predicates in by-operator variation. This is illustrated in Figure~\ref{fig:op-pred-ratings}, which shows mean projection ratings for the 20 attitude predicates by embedding operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
			%
			For instance, whereas the CC of \emph{be annoyed} projects more from under negation and questions than conditionals and modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals. The CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant.
			%
			These generalizations are supported by models\ \# 2--4 in Table\ \ref{t:models}, which also each have at least $34$ highly significant interaction terms (out of $57$ possible interactions of operator and predicate).

			\begin{figure}[ht]
				\centering
				\includegraphics[width = \linewidth]{projective-pred-op}
				\caption{Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals. Embedding operator coded by letter and color:  \texttt{N} (light blue): negation, \texttt{M} (orange): modals, \texttt{C} (green): conditional antecedents, \texttt{Q} (dark blue): polar questions.}
				\label{fig:op-pred-ratings}
			\end{figure}

@@

			This result that by-operator projection variability interacts with predicate concurs with \citet{smith_relationship_2014}, while we did not reproduce their result that the CC of \emph{know} projects more from negation than conditionals. This differs from the result of \citet{sieker_projective_2022}, who found no interaction between predicates and operators.
			%
			Given that the methods and the set of projective contents of our experiment were very similar to theirs, this difference may suggest cross-linguistic variation in projection variability (cf. \citealt{tonhauser_projection_2020}).

		\subsubsection{Gradience and lexical classes}

			Finally, our results provide further support (from negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates, while also replicating it for polar questions. 
			E.g., the CCs of non-factive \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.

			Similarly, our results question \posscite{karttunen_observations_1971} proposed difference between factive and semi-factive predicates, in line with \citet{sieker_projective_2022} (see also \citealt{beaver_have_2010}): Although our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of semi-factive \emph{discover} is less projective than that of \emph{know}, the same does not hold in conditionals.
			%
			Further, contrary to what would be expected based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates, the CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC of \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators. The pattern observed for {\em know} does not fit into either category.

		\subsubsection{Converging evidence for operator / predicate interactions}
			MegaVeridicality dataset (\citealt{white_role_2018}): 517 predicates in three sentence types

			\ex. \a. Somebody didn’t know that a particular thing happened. (Did that thing happen?)
				\b. If somebody knows that a particular thing happened, did that thing happen?
				\b. If somebody didn’t know that a particular thing happened, did that thing happen?
				\z.
			\z.

			\begin{figure}[ht]
				\centering
				\includegraphics[width = \linewidth]{mega-veridicality}
				\caption{Mean projection ratings by embedding context and predicate.}
				\label{fig:figure1}
			\end{figure}

@@

			Results:

			\begin{enumerate}
				\item Effect of operator on projection

				\item Effect of operator on projection differs by predicate

				\item By-predicate variation on operator effects are gradient and cannot be predicted by categorical lexical classes

			\end{enumerate}

			Methodological implications:
			\begin{itemize}
				\item projection ratings do not categorically distinguish between lexical classes (e.g., factives, semi-factives, veridical predicates), so future research appealing to these categories must clarify their definition.

				\item Additionally, claims about projection variability must be relativized to the entailment-canceling operator.

				\item but we can still use FoS test in teaching, JT made this claim, but I don't understand the argument... help?

			\end{itemize}

			Theoretical implications: Accounts of projective content need to take into account predicate/operator interactions, and the gradient by-predicate variability. The following section argues that our results provide evidence for a triggering algorithm based on contextual rather than lexical or conceptual information (e.g., in the sense of \citealt{abrusan_predicting_2011,simons_best_2017}), and that the input to this triggering algorithm should be contextual rather than lexical inferences (e.g., in the sense of \citealt{schlenker_triggering_2021}).

\section{Theoretical implications}
	Our results---that projection is modulated by entailment-canceling operators, that there is by-predicate variation in the effect of operator on projection, and that this variation is gradient in ways that cannot be predicted by categorical lexical classes---are not captured by contemporary projection analyses, for several reasons.

	Here, we discuss (i) dynamic approaches, which assume that projection is determined lexically for factive predicates (\citealt{heim_projection_1983,heim_presupposition_1992,van_der_sandt_presupposition_1992}), (ii) approaches which assume that projection is determined by a contextual triggering algorithm that operates over the lexical entailments of an utterance (\citealt{abrusan_predicting_2011,simons_best_2017}), and (iii) approaches with a conceptual triggering algorithm that operates over contextual entailments of an utterance (\citealt{schlenker_triggering_2021}). While all of these approaches meet limitations in accounting for our data, we argue that the observed gradient effects of predicate/operator interactions favor integrating insights from competing accounts, namely a contextual triggering algorithm (as in \citealt{abrusan_predicting_2011,simons_best_2017}) operating over contextual set of inferences (as in \citealt{schlenker_triggering_2021}).

	- effect of different entailment-cancelling operators, not captured (heim vs santd, if projects projects; abtrusan operators?; schlenker can be imagined) in all systems it is conceivable, but no systemaitc itneraction spelled out

	- rely on lexical classes, factives, veriudical, no predictions of non-veridicals

	- not capture how the operator differentially affects different predicates (contextual entailment may provied an approach, but also needs to be spelled out yet)

>>>>

{\large \textbf{Projection variability of clausal complements across different operators}}\footnote{We would like to thank\ldots }\\
Lisa HOFMANN --- \textit{University of Stuttgart}\\
Marie-Catherine DE MARNEFFE --- \textit{UC Louvain}\\
Judith TONHAUSER --- \textit{University of Stuttgart}\\

\textbf{Abstract.}
	We present experimental evidence that the projection of clausal complements \textbf{(i)} varies among entailment-canceling operators, \textbf{(ii)} that the effect of operator varies between clause-embedding predicates, and \textbf{(iii)} we extend the result from \citet{degen_are_2022}, that projection ratings in polar questions are gradient in ways that cannot be predicted by categorical lexical classes, to contexts with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
	%
	The observed variability is not captured by existing theoretical accounts of projection
	(e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}).
	%
	Our results suggest that an analysis must consider interactions between predicates and operators and raise important questions for future research on projection.

@@

\section{Introduction}
	Language users may infer that a speaker who uses a clause-embedding predicate, as in \ref{ex:family}, is committed to the content of the complement (CC, here: \emph{Julian dances salsa,}) even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditional antecedents \ref{ex:cond}, in which case we say that it \textit{projects} (e.g., \citealt{frege_uber_1892,strawson_referring_1950,kiparsky_fact_1970,karttunen_conventional_1979}).

@@

	We investigate whether the projection of embedded propositional content from under entailment-cancelling operators differs by operator and whether by-operator differences vary between embedding predicates.
	%
	Textbook diagnostics for projection often assume that if a content projects across one of the operators in \Last, it uniformly projects across the others (e.g., \citealt{chierchia_meaning_1990}). Accordingly, current research examines projection variability between these operators only rarely and with conflicting findings.

@@

	\citet{smith_relationship_2014} investigated by-operator variation, comparing negation and conditionals, for various types of projective content. They found that the expressive content of epithets (e.g. \textit{idiot}) and the CC of \textit{know} was more projective under negation than conditionals, whereas appositive relative clauses and the preparatory content of \textit{win} showed the opposite pattern, and the existential presupposition of clefts showed no difference. However, their task involved asking participants how surprised they would be to learn the content under investigation after observing the utterance. Since other factors besides speaker commitment modulate discourse (un)expectedness \citep[see e.g.][]{zimmermann_grammatical_2011,tonnis_german_2021}, it is unclear that they (only) measured projection.

	\posscite{sieker_projective_2022} study on German clause-embedding predicates compared the projection of purported factives (\textit{wissen} `know', \textit{bereuen} `regret', \textit{enthüllen} `reveal') and semi-factives (\textit{bemerken} `notice', \textit{entdecken} `discover', \textit{herausfinden} `find out') across the four contexts in \Last. They replicated Smith and Hall's result that the CC of \textit{know} projects more from under negation than conditionals for German \textit{wissen} (`know'). However, this finding was part of an overall pattern of higher projection ratings with negation than other operators.
	%
	Importantly, their analysis did not explore the impact of operators on each predicate individually. Instead, it grouped ratings for the four purported factive and semi-factive predicates, respectively. Based on this analysis, they found no interaction of operator with predicate type, and thus no evidence for the distinction of factives vs. semi-factives.

	Studies in \citet{djarv_cognitive_2018}, \citet{tonhauser_how_2018}, and \citet{degen_are_2022} observed by-predicate variation in polar questions, where Karttunen's classification predicts more projection for factives than semi-factives.
	%
	\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, did find higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive). But here, it is also not obvious how exactly this task relates to projection.\footnote{Say why}

	\citet{tonhauser_how_2018} and \citet{degen_are_2022} measured speaker commitments towards a broad range of projective contents more directly, by collecting judgments of how certain participants take a speaker to be of the contents under investigation.
	%
	The observed differences between CCs of various clause-embedding predicates do not align with theoretically assumed lexical classes. While \citet{degen_are_2022} demonstrate this for the distinction between purported factive and non-factive predicates, the results did not match the expectations from Karttunen's factive vs. semi-factive classification either.
	%
	\citet{tonhauser_how_2018} found, e.g., that the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}. \citet{degen_are_2022} observed that the CC of factive \textit{reveal} was less projective than that of semi-factive \textit{discover}, which, in turn, was less projective than that of factive \textit{know}. These findings suggest that examining projection ratings for various predicates individually, rather than pre-grouping them based on theoretical assumptions, can provide more nuanced insights into by-predicate projection variation.

	While \citet{smith_relationship_2014} found by-predicate variation in the effect of the operator on projection, \citet{sieker_projective_2022} did not.
	%
	These divergent results raise the question if they are due to cross-linguistic variation, task differences, or artifacts of the analysis. To address this, we conducted a series of experiments designed to assess projection across the four entailment-canceling operators in \ref{ex:family}. We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive and semi-factive predicates, and 15 non-factive predicates, given recent findings that their complements may also project, albeit to varying degrees (\citealt{degen_are_2022}).

@@

	To assess the effect of entailment-cancelling operators and clause-embedding predicates on the projection of the embedded content, we collected projectivity judgments for contents embedded under 20 clause-embedding predicates in four sets of experiments. The predicates were embedded under negation in Experiments~1, under polar questions (Exps.~2), under the epistemic possibility modal {\em perhaps} (Exps.~3), and in conditional antecedents (Exps.~4). Each set of experiments contained three experiments differing in an at-issueness measure used in a separate block. We focus on the projection ratings here.

		Projection was measured across all 12 experiments with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,lorson_influence_2018,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}; and discussion in \citealt{tonhauser_how_2018}.} Participants were presented with utterances like those in \Next and judged whether the speaker (who was named) was certain of the CC (e.g.: Is Christopher certain that Julian dances salsa?).

		\ex. \label{ex:certain-that}
			\a. Christopher: \emph{\lq Cole didn't discover that Julian dances salsa.\rq}
			\b. Christopher: \emph{\lq Did Cole discover that Julian dances salsa?\rq}
			\c. Christopher: \emph{\lq Perhaps Cole discovered that Julian dances salsa.\rq}
			\d. Christopher: \emph{\lq If Cole discovered that Julian dances salsa, Logan will be joyful.\rq}
			\z.
			Projection question: Is Christopher certain that Julian dances salsa?
		\z.

		Following \citet{tonhauser_prosodic_2016,djarv_prosodic_2017,tonhauser_how_2018,lorson_influence_2018,mahler_does_2019}, we took a `yes' response to indicate that participants interpret the utterance in a way that the presented speaker (e.g., Christopher in \Last) is committed to the embedded content, which reflects projection. A `no' response was taken to indicate that the content does not project.
		%
		Based on Karttunen's generalization, we would expect the CC of semi-factive predicates to receive high projection ratings under negation and lower ratings under the other operators, while the CC of factive predicates would consistently receive relatively high projection ratings throughout.

@@

		\subsubsection{Participants}
			In each of the 12 experiments, 250-300 inhabitants of the USA participated via Prolific or Amazon's MT platform. 
			%
			We analyze the data from $2,682$ self-reported native speakers of American English, after excluding participants who failed one of the following two attention-checks: their mean rating on six control items is more than 2 sd different from the group mean, or they always roughly selected the same point on the scale for target stimuli (identified by low variance, and then manually inspecting their response patterns).

		\subsubsection{Materials and procedure }
			We tested 20 clause-embedding predicates \Next which have shown projection variability in question contexts (\citealt{degen_are_2022}). They comprise three purported factive predicates \Next[a], two purported semi-factives \Next[b], nine that have been characterized as optionally factive in \citealt{kiparsky_fact_1970} \Next[c], and six non-factives which include two veridical predicates \Next[d], and four non-veridical ones \Next[e].

			\ex. 20 clause-embedding predicates \hfill \citealt{degen_are_2022}
				\a. canonically factive:
					\hfill \emph{be annoyed, know, reveal}
				\b. semi-factive:
					\hfill \emph{discover, see}
				\b. optionally factive: \newline
					\phantom.\hfill \emph{acknowledge admit, announce, confess, confirm, establish, hear, inform, prove}
				\b. Non-factive veridical:
					\hfill \emph{be right, demonstrate}
				\b. Non-veridical:
					\hfill \emph{pretend, say, suggest, think}
				\z.
			\z.

			Each participant saw all 20 predicates, randomly paired with a unique embedded clause from a set of 20 contents, and a subject from a set of 20 proper names. The predicates associated with projective content were embedded under (i) negation, (ii) polar questions, (iii) the epistemic possibility modal \textit{perhaps}, and (iv) in conditional antecedents in the four different sets of experiments.

			On each trial, participants read an utterance and the corresponding projection question. They gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other. This is illustrated in Figure~\ref{fig:trial}. Following, e.g., \citet{tonhauser_how_2018}, we take high ratings of speaker certainty to reflect speaker greater speaker commitment towards the CC.

@@

			Each experiment also included six control stimuli, which were main clauses with appositive relative clauses, which were also presented as utterances made by a named speaker. Here, the main clause content was targeted by the projection question. In the experiments with polar questions, the controls were expected to receive low certainty ratings, whereas in the experiments with declaratives, they were expected to receive high ratings.

			Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody say something to somebody else. Each participant saw a random set of 26 stimuli: each set contained one target utterance for each of the 20 contents (each paired with a unique clause-embedding predicates) and the same 6 (experiment-specific) control stimuli. Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block. Block order and within-block trial order was randomized. After completing the experiment, participants filled out a short optional demographic survey.

	\subsection{Results and discussion}
		Our data reveals three key results: \textbf{(i)} There is projection variability by operator; \textbf{(ii)} There is by-predicate variation in the effect of operator on projection; and \textbf{(iii)} We find further support (from entailment-cancelling operators beyond polar questions) for \posscite{degen_are_2022} result that by-predicate projection variation is gradient in a way that cannot be predicted by categorical lexical classes.

		\subsubsection{By-operator variation}
			When aggregating across predicates, we found an effect of \texttt{operator}. The distribution of projection ratings, as well as means and $95\%$ bootstrapped confidence intervals by operator are shown in Figure~\ref{fig:op-ratings}.

			\begin{figure}[ht]
				\centering
				\includegraphics[scale = .8]{operator-graph-1}
				\caption{Certainty ratings by operator with means and $95\%$ bootstrapped confidence intervals.}
				\label{fig:op-ratings}
			\end{figure}

			Mean projection ratings were higher for CCs in question-embeddings than under negation and modals, but lower than in conditional antecedents. These generalizations are supported by Model \#1 reported in Table~\ref{t:op-model}.

@@

					\caption{Excerpt from the model output of a linear mixed effects model with fixed effects of operator; random effects: participant and item intercepts, fitted with \texttt{lme4, lmertest} in \texttt{R}.\label{t:op-model}}
				\end{table}

			Like \citet{sieker_projective_2022}, we observe a small main effect of entailment-cancelling operator on projection. However, our result, that the CC of clause-embedding predicates projects most from conditional antecedents and least from under negation and modals, differs from their finding that the CC of the German predicates they tested projects more from under negation than the other operators. Further, the projection ratings they report are overall quite a bit higher than the ones reported here (their overall mean is at around $0.855$, while ours is at around $0.501$).
			%
			We can understand these differences based on an effect of \texttt{predicate}, established in the upcoming subsections, and a different set of predicates tested in the two studies. While Sieker \& Solstad tested only predicates that are characterized as factive or semi-factive in the literature, yielding relatively high projection ratings, we also included ones that are usually characterized as non-factive, lowering the overall mean projection ratings. The difference in the direction of the effect of the operator between the two studies can also be tied to a different set of predicates tested, because the effect of operator on projection differs by predicate.

		\subsubsection{Gradient by-predicate variation}

			Taking into account the effect of \texttt{predicate} on projection, we find gradient projection variation, like \citet{tonhauser_how_2018} and \citet{degen_are_2022}.
			%
			This is illustrated in Figure~\ref{fig:op-pred-ratings}, which shows mean projection ratings for the 20 predicates grouped by embedding operator. Here, predicates on the x-axis are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).

			\begin{figure}[ht]
				\centering
				\includegraphics[width = \linewidth]{predicate-operator-graph-1}
				\caption{Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals. Embedding operator coded by letter and color:  \texttt{N} (light blue): negation, \texttt{M} (orange): modals, \texttt{C} (green): conditional antecedents, \texttt{Q} (dark blue): polar questions.}
				\label{fig:op-pred-ratings}
			\end{figure}

			Our results therefore provide further support from negation, modals, and conditionals for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates, while also replicating it for polar questions. The CCs of non-factive \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.

@@

			We observe differences between predicates in by-operator variation, i.e., a operator/predicate interaction effect, which is also illustrated in Figure~\ref{fig:op-pred-ratings}.
			%
			For instance, whereas the CC of \emph{be annoyed} projects more from under negation and questions than conditionals and modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals. The CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant.
			%
			These generalizations are supported by models\ \# 2--4 in Table\ \ref{t:models}, which also each have at least $34$ highly significant interaction terms (out of $57$ possible interactions of operator and predicate).

@@

			Our result that by-operator projection variability interacts with predicate concurs with \citet{smith_relationship_2014}, while we did not reproduce their result that the CC of \emph{know} projects more from negation than conditionals (we found no difference here). Our results differ from those of \citet{sieker_projective_2022}, who found no difference between the two groups of predicates they assumed in the effect of operator.
			%
			However in line with Sieker \& Solstad, our results question \posscite{karttunen_observations_1971} proposed difference between factive and semi-factive predicates (see also \citealt{beaver_have_2010}). Although our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of semi-factive \emph{discover} is less projective than that of \emph{know}, the same does not hold in conditionals, contrary to what would be expected based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates.
			%
			Further, the CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC of \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators. The pattern observed for {\em know} does not fit into either category.

		\subsubsection{Converging evidence for operator / predicate interactions}

			Our experimental result that that the effect of entailment-cancelling operator differs by predicate finds further support in the data from \posscite{white_role_2018} MegaVeridicality dataset.
			%
			White \& Rawlins assess projection inferences associated with 517 clause-embedding predicates by presenting them in 50 \lq low-content\rq\ syntactic frames, like the ones in \Next, illustrated here with \textit{know}. The frames in \Next present predicates associated with projective contents embedded under negation \Next[a], in conditional antecedents \Next[b], or both \Next[c], in combination with a projection question (\textit{Did that thing happen?}).

			\ex. \a. Somebody didn’t know that a particular thing happened. Did that thing happen?
				\b. If somebody knows that a particular thing happened, did that thing happen?
				\b. If somebody didn’t know that a particular thing happened, did that thing happen?
				\z.
			\z.

			In a series of experiments, participants presented with stimuli like \Last responded with \lq yes\rq, \lq no\rq, or  \lq maybe\rq (coded as \texttt{1, -1,} or \texttt{0}, respectively). Here, we present data from their data set, for the frames in \Last, and 16 of their predicates. These include ones also used in our experiment as purported factives (\textit{be annoyed, know, reveal}) and semi-factives (\textit{discover, see}), and eleven further predicates commonly characterized as factive or semi-factive (\textit{amuse, find out, forget, learn, love, notice, realize, recognize, regret, remember, understand}).

			The mean projection ratings by embedding context and predicate are presented in Figure~\ref{fig:figure4}, which shows that effect of embedding context on projection differs by predicate.

			For instance,\footnote{Give some examples of differences here, but first double check if lines are labeled correctly. Also angle labels on x-axis.}

			% the predicates \textit{see} and \textit{learn} have the same mean ratings for both negative and conditional contexts. However, the effect of negative conditional contexts does not simply reflect the additive result of both effects, but

			\begin{figure}[ht]
				\centering
				\includegraphics[width = \linewidth]{mega-veridicality}
				\caption{Mean projection ratings by embedding context and predicate.}
				\label{fig:figure4}
			\end{figure}

			It is worth noting that the task assesses global projection only in case of \Last[a], but not \Last[b+c], where the question \textit{did it happen?} is embedded in the conditional consequent. It assesses whether the projective inference holds in the conditional local context. Nevertheless, the data provides supporting evidence for by-predicate variation in the effect of operators on projection. Crucially, the by-predicate variation in the effect of negation in the global context is different from the by-predicate variation in the effect of negation in the conditional context.

@@

			Based on our data, we conclude that embedded propositional content projects differently from under different entailment-cancelling operators. A methodological implication of this finding is that claims about projection variability must be relativized to the entailment-canceling operator in future research.
			%
			Further, the observed by-predicate variation does not categorically distinguish between lexical classes (e.g., factives, semi-factives, veridical predicates), so future research appealing to these categories must clarify their definition.
			%
			Finally, the effect of operator on projection differs by predicate. The theoretical implications of our findings are discussed in the following section. Importantly, accounts of projective content need to take into account predicate/operator interactions, and the gradient by-predicate variability.

\section{Theoretical implications}
	Our results---that projection is modulated by entailment-canceling operators, that there is by-predicate variation in the effect of operator on projection, and that this variation is gradient in ways that cannot be predicted by categorical lexical classes---are not captured by contemporary projection analyses.

		Here, we discuss this for (i) dynamic approaches, which assume that projection is determined lexically for factive predicates (\citealt{heim_projection_1983,heim_presupposition_1992,van_der_sandt_presupposition_1992}), (ii) approaches which assume that projection is determined by a triggering algorithm based on information structure that operates over the literal entailments of an utterance (\citealt{abrusan_predicting_2011,simons_best_2017}), and (iii) approaches with a triggering algorithm based on world-knowledge that operates over contextual entailments of an utterance (\citealt{schlenker_triggering_2021}). While all of these approaches meet limitations in accounting for our data, we argue that the observed gradient effects of predicate/operator interactions favor integrating insights from competing accounts, namely a contextual triggering algorithm (as in \citealt{abrusan_predicting_2011,simons_best_2017}) operating over contextual set of inferences (as in \citealt{schlenker_triggering_2021}).

	\subsection{Lexical triggering in dynamic semantics}

		Dynamic semantic accounts of projection (e.g., \citealt{heim_projection_1983,heim_presupposition_1992,van_der_sandt_presupposition_1992}) assume lexical triggering: there is a class of (semi-)factive predicates, which lexically presuppose their CCs. Presupposed CCs project to the global context, except when that would produce an inconsistency, in which case they are accommodated to the local context of the operator.
